### 2014-10-22 mental-model-paper/talk-questionnaire.tex
questionnaire work

\documentclass{article}
\usepackage{lipsum}
\newcounter{question}
\setcounter{question}{0}
\begin{document}

\newcommand\Que[1]{%
   \leavevmode\par
   \stepcounter{question}
   \noindent
   \thequestion. Q --- #1\par}

\newcommand\Ans[2][]{%
    \leavevmode\par\noindent
   {\leftskip37pt
    A --- \textbf{#1}#2\par}}

\par
\noindent
\textit{Rachel Rivera}

\Que{Give at least two specific examples of objects or behaviors that cognitive psychologists study.}
\Ans{ \\ 1. The behaviors of patients with localized brain injury. \\ 2. The behaviors of split-brain patients.}

\Que{How do these subjects of study relate to what interaction design studies?}
\Ans{These subjects relate to a plethora of different interaction design topics. One prominent topic that is relevant in both fields is the notion of \textit{learnability}. }

\Que{Give at least two specific \textit{methods} used by cognitive psychologists to study or learn about their subjects.}
\Ans{\\ 1. Having split-brain patients draw the same object twice--- once with the left hand and once with the right. \\ 2. Examining what each hemisphere can accomplish in isolation by testing split-brain subjects' responses to touching objects versus saying the names of the objects. }

\Que{How do these methods of study relate to how interaction design research is performed?}
\Ans{ The manner in which these methods isolate portions of the brain and examine the abilities of those portions in isolation parallels the general manner in which interaction design research is performed. }
\end{document}